---
title: "McKibben DSC520 Ex. 6.2"
author: "Makayla McKibben"
date: "2024-07-12"
output: pdf_document
---

# Iris Analysis 
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)
new_iris <- data.frame(iris)
```

## Average Sepal Length by Species

```{r average sepal length, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
head(str(new_iris))
species_s1 <- select(new_iris, Sepal.Length, Species)

species_1 <- slice(species_s1,1:50)
setosa_avg_sl <- mean(species_1$Sepal.Length)

species_2 <- slice(species_s1,51:100)
versicolor_avg_sl <- mean(species_2$Sepal.Length)

species_3 <- slice(species_s1,101:150)
virginica_avg_sl <- mean(species_3$Sepal.Length)

setosa_avg_sl
versicolor_avg_sl
virginica_avg_sl
```
```{r iris data splitting, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
species_sw <- select(new_iris, Sepal.Width, Species)

species_1_sw <- slice(species_sw,1:50)
species_2_sw <- slice(species_sw,51:100)
species_3_sw <- slice(species_sw,101:150)

species_pl <- select(new_iris, Petal.Length, Species)

species_1_pl <- slice(species_pl,1:50)
species_2_pl <- slice(species_pl,51:100)
species_3_pl <- slice(species_pl,101:150)

species_pw <- select(new_iris, Petal.Width, Species)

species_1_pw <- slice(species_pw,1:50)
species_2_pw <- slice(species_pw,51:100)
species_3_pw <- slice(species_pw,101:150)
```

```{r ggplot2 load, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```


## Visualizations
```{r plots, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
setosa_hist_sl <- ggplot(species_1, aes(Sepal.Length))
setosa_hist_sl + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Setosa Sepal Length", x ="Sepal Length", 
       y = "Frequency") + xlim(4, 6)
versicolor_hist_sl <- ggplot(species_2, aes(Sepal.Length))
versicolor_hist_sl + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Versicolor Sepal Length", x ="Sepal Length", 
       y = "Frequency") + xlim(4.5, 7.5)
virginica_hist_sl <- ggplot(species_3, aes(Sepal.Length))
virginica_hist_sl + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Virginica Sepal Length", x ="Sepal Length", 
       y = "Frequency") + xlim(4.5, 8)

setosa_hist_sw <- ggplot(species_1_sw, aes(Sepal.Width))
setosa_hist_sw + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Setosa Sepal Width", x ="Sepal Width", 
       y = "Frequency") + xlim(2, 4.5)
versicolor_hist_sw <- ggplot(species_2_sw, aes(Sepal.Width))
versicolor_hist_sw + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Versicolor Sepal Width", x ="Sepal Width", 
       y = "Frequency") + xlim(2, 4)
virginica_hist_sw <- ggplot(species_3_sw, aes(Sepal.Width))
virginica_hist_sw + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Virginica Sepal Width", x ="Sepal Width", 
       y = "Frequency") + xlim(2, 4)

setosa_hist_pl <- ggplot(species_1_pl, aes(Petal.Length))
setosa_hist_pl + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Setosa Petal Length", x ="Petal Length", 
       y = "Frequency") + xlim(0.5, 2)
versicolor_hist_pl <- ggplot(species_2_pl, aes(Petal.Length))
versicolor_hist_pl + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Versicolor Petal Length", x ="Petal Length", 
       y = "Frequency") + xlim(2.5, 5.5)
virginica_hist_pl <- ggplot(species_3_pl, aes(Petal.Length))
virginica_hist_pl + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Virginica Petal Length", x ="Petal Length", 
       y = "Frequency") + xlim(4, 7)

setosa_hist_pw <- ggplot(species_1_pw, aes(Petal.Width))
setosa_hist_pw + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Setosa Petal Width", x ="Petal Width", 
       y = "Frequency") + xlim(0, .75)
versicolor_hist_pw <- ggplot(species_2_pw, aes(Petal.Width))
versicolor_hist_pw + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Versicolor Petal Width", x ="Petal Width", 
       y = "Frequency") + xlim(0.75, 2) 
virginica_hist_pw <- ggplot(species_3_pw, aes(Petal.Width))
virginica_hist_pw + geom_histogram(binwidth = 0.05, fill = "lightgrey", color = "black") + 
  labs(title = "Virginica Petal Width", x ="Petal Width", 
       y = "Frequency") + xlim(1.25, 2.75)
setosa_box_sl <- ggplot(species_1, aes(Sepal.Length))
setosa_box_sl + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Setosa Sepal Length", x ="Sepal Length", 
       y = "Frequency")
versicolor_box_sl <- ggplot(species_2, aes(Sepal.Length))
versicolor_box_sl + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Versicolor Sepal Length", x ="Sepal Length", 
       y = "Frequency") 
virginica_box_sl <- ggplot(species_3, aes(Sepal.Length))
virginica_box_sl + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Virginica Sepal Length", x ="Sepal Length", 
       y = "Frequency")

setosa_box_sw <- ggplot(species_1_sw, aes(Sepal.Width))
setosa_box_sw + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Setosa Sepal Width", x ="Sepal Width", 
       y = "Frequency")
versicolor_box_sw <- ggplot(species_2_sw, aes(Sepal.Width))
versicolor_box_sw + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Versicolor Sepal Width", x ="Sepal Width", 
       y = "Frequency") 
virginica_box_sw <- ggplot(species_3_sw, aes(Sepal.Width))
virginica_box_sw + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Virginica Sepal Width", x ="Sepal Width", 
       y = "Frequency")

setosa_box_pl <- ggplot(species_1_pl, aes(Petal.Length))
setosa_box_pl + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Setosa Petal Length", x ="Petal Length", 
       y = "Frequency")
versicolor_box_pl <- ggplot(species_2_pl, aes(Petal.Length))
versicolor_box_pl + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Versicolor Petal Length", x ="Petal Length", 
       y = "Frequency") 
virginica_box_pl <- ggplot(species_3_pl, aes(Petal.Length))
virginica_box_pl + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Virginica Petal Length", x ="Petal Length", 
       y = "Frequency")

setosa_box_pw <- ggplot(species_1_pw, aes(Petal.Width))
setosa_box_pw + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Setosa Petal Width", x ="Petal Width", 
       y = "Frequency")
versicolor_box_pw <- ggplot(species_2_pw, aes(Petal.Width))
versicolor_box_pw + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Versicolor Petal Width", x ="Petal Width", 
       y = "Frequency") 
virginica_box_pw <- ggplot(species_3_pw, aes(Petal.Width))
virginica_box_pw + geom_boxplot(fill = "lightblue", color = "black", notch = TRUE) + 
  labs(title = "Virginica Petal Width", x ="Petal Width", 
       y = "Frequency")
```
Iris species Setosa has the shortest average sepal length of this data set while Virginica has the longest. Versicolor is in between these two.
Average sepal width of Versicolor is the shortest, with Virginica in the middle, and Setosa being the longest.
The average petal length increases from Setosa, the shortest, with Versicolor in the middle and Virginica as the longest.
Average petal width increases from Setosa, the shortest, with Versicolor in the middle and Virginica being the longest.