---
title: "McKibben_Assingment_8.2"
author: "Makayla McKibben"
date: "2024-07-29"
output: pdf_document
---

```{r setup, include = TRUE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("purrr")
library(purrr)
#install.packages("tidyverse")
library(stringr)
library(dplyr)
#install.packages("Metrics")
library(car)
library(Metrics)
```

# Call and Clean Housing Data
```{r call and clean, echo = TRUE, include = TRUE}

housing <- read.csv(file = 'Housing.csv', header = TRUE, sep =",", stringsAsFactors = FALSE)
head(housing, 5)
mod_sq_ft <- cbind(housing$Sale.Price, housing$sq_ft_lot)
colnames(mod_sq_ft) <- c("Sales Price", "Lot sq. ft.")
head(mod_sq_ft, 5)
mod_sq_ft <- as.data.frame(mod_sq_ft)
class(mod_sq_ft)
head(mod_sq_ft, 5)
mod_sq_ft <- mod_sq_ft[complete.cases(mod_sq_ft),]
class(mod_sq_ft)
head(mod_sq_ft, 5)
plot(mod_sq_ft$`Lot sq. ft.`, mod_sq_ft$`Sales Price`)
model_sq_ft <- lm(`Sales Price` ~ `Lot sq. ft.`, data = mod_sq_ft)
summary(model_sq_ft)
```
# Process
Take a quick look to see how the data is structured
Group the needed data
Assign relevant headers
Check structure of new data set
Turn it into a data frame
Check structure of new data set
Remove empty rows
Check structure of new data set
Plot data to get an idea of the relationship
Create model
Check values

# Explain Results
R2 is 0.01435 and adj. R2 is 0.01428, this converts to an accuracy of 1% for prediction of sales price from lot square footage.
The intercept is 641800 with a standard error of 3800. The slope is 0.851 with a std. error of 0.0622.
The F-ratio is 187.3 with a p-value of 2.2e-16 so we should consider that it is significant.


```{r first residuals, echo = TRUE, include = TRUE}
# Get residuals
mod_sq_ft$residuals <- resid(model_sq_ft)
mod_sq_ft$std_residuals <- rstandard(model_sq_ft)
mod_sq_ft$stud_residuals <- rstudent(model_sq_ft)

# Plot residuals
plot(model_sq_ft)
hist(mod_sq_ft$std_residuals)
qqPlot(mod_sq_ft$residuals)
```
# Explain Results
The Residuals vs. Fitted data is not an even distribution. There is a group that is more concentrated and overall it makes the general shape of a triangle indicating Heteroscedasticity.

# Explain QQ Plot
The qq plot shows the residuals form a sigmoid shape that deviates from the central line, this indicates that the residuals are not normally distributed. This is verified by a histogram of the standardized residuals.

```{r second model prep, echo = TRUE, include = TRUE}
bed_bath <- housing$bedrooms + housing$bath_full_count + housing$bath_half_count + housing$bath_3qtr_count
mod_mp <- cbind(housing$Sale.Price, housing$square_feet_total_living, bed_bath)
colnames(mod_mp) <- c("Sales Price", "Sq. ft. Total Living", "Bedrooms + Bathrooms")
head(mod_mp, 5)
mod_mp <- as.data.frame(mod_mp)
class(mod_mp)
head(mod_mp, 5)
mod_mp <- mod_mp[complete.cases(mod_mp),]
class(mod_mp)
head(mod_mp, 5)
plot(mod_mp$`Sq. ft. Total Living`, mod_mp$`Sales Price`)
plot(mod_mp$`Bedrooms + Bathrooms`, mod_mp$`Sales Price`)
model_mp <- lm(`Sales Price` ~ `Sq. ft. Total Living` + `Bedrooms + Bathrooms`, data = mod_mp)
summary(model_mp)
```

# Second Model Process
Create new variable that combines the total number of bedrooms and bathrooms
Group the needed data
Assign relevant headers
Check structure of new data set
Turn it into a data frame
Check structure of new data set
Remove empty rows
Check structure of new data set
Plot data to get an idea of the relationship
  Linear, positively correlated
Plot data to get an idea of the relationship
  Linear, positively correlated
Create model
Check values

# Variable Choice
I believe that liveable square feet and the number of bedrooms and bathrooms are going to be better indicators of sales price than the size of the lot the house is on.
People often have requirements for the number of bedrooms and bathrooms a home may have, this is often more important than the size of the lot.

# Explain Model Results
The R2 value is .2072 and the adjusted R2 is 0.2071, this indicates that our model predicts 20.7% of our sales price value from the predictor variables. That is nearly 20% better than the first model.
The intercept is 222,319 with a std. error of 14,106. The b1 (Sq. ft. of the living space) value is 196 with a std. error of 4.8 and b2 (Bedrooms + Bathrooms) is -9578 with a std. error  of 3192.
The F-ratio is 10x larger than in our previous model and the p-value is < 0.001 so there is a 99.9% chance we reject the null hypothesis.

```{r second model residuals, echo = TRUE, include = TRUE}
# Obtain residuals of new model
mod_mp$residuals <- resid(model_mp)
mod_mp$std_residuals <- rstandard(model_mp)
mod_mp$stud_residuals <- rstudent(model_mp)

# Plot residuals
plot(model_mp)
hist(mod_mp$std_residuals)
qqPlot(mod_mp$residuals)
```

# Explain the Residuals
The Residuals vs. Fitted data is not an even distribution although, it is closer than the previous model. There is a group that is more concentrated and overall it makes the general shape of a triangle indicating Heteroscedasticity.

# Explain QQ Plot
The qq plot shows the residuals form a sigmoid shape that deviates from the central line, this indicates that the residuals are not normally distributed. This is verified by checking a histogram of the standardized residuals.

```{r model comparison, echo = TRUE, include = TRUE}
anova(model_sq_ft, model_mp)
```
# Anova Comparison
The anova test gave us an F(1, 12862) = 3128 with a Pr(>F) of < 2.2e-16 which tells us the second model is a significant improvement over the original model.

# Bias Comparison
I believe both models are biased similarly based on comparisons of the residual v fitted and qq plots. The linear fit of both indicate that the low end is biased high and the high end is biased low.

```{r model comparison RMSE, echo = TRUE, include = TRUE}
# Prediction accuracy first model
lot_preds <- predict(model_sq_ft, newdata = mod_sq_ft)
lot_rmse <- rmse(mod_sq_ft$`Sales Price`, lot_preds)
lot_rmse

# Prediction accuracy second model
mp_preds <- predict(model_mp, newdata = mod_mp)
mp_rmse <- rmse(mod_mp$`Sales Price`, mp_preds)
mp_rmse

# RMSE Difference
diff_rmse <- lot_rmse - mp_rmse
diff_rmse
```
# RMSE Comparison
## RMSE for the First Model
401,452.5

## RMSE for Second Model
360,043.1

Yes, the second model improved upon the RMSE relative to the first model. The difference is 41409.44