---
title: "McKibben_DSC520_Ex_7.2"
author: "Makayla McKibben"
date: "2024-07-28"
output: pdf_document
---

```{r plots, echo = TRUE, include = TRUE}
# Install appropriate packages
# install.packages("tidyverse")

# Importing the data set
survey <- read.csv(file = 'student-survey.csv', header = TRUE, sep =",", stringsAsFactors = FALSE)

# Looking at a section of the data set to understand the structure
head(survey)

# Calling relevant library
library(ggplot2)

# Creating plots
r_v_tv <- ggplot(survey, aes(TimeReading, TimeTV))
r_v_tv + geom_point(color = "navy", shape = 8, size = 4.8) +
  theme(panel.grid = element_line(color = "lightgrey", linewidth = 0.8, linetype = 1), 
        panel.background = element_rect(color = "white", fill = "darkgrey")) +
  labs(title = "Time Spent", x ="Time Reading", 
       y = "Time TV") + xlim(0,7)

happy_reader <-ggplot(survey, aes(TimeReading, Happiness))
happy_reader + geom_point(color = "purple", shape = 8, size = 4.8) + 
  theme(panel.grid = element_line(color = "lightgrey", linewidth = 0.8, linetype = 1),
        panel.background = element_rect(color = "white", fill = "darkgrey")) +
  labs(title = "Happy Readers", x ="Time Reading", 
       y = "Happiness") + xlim(0,7) + ylim(45,95)

happy_tv <-ggplot(survey, aes(TimeTV, Happiness))
happy_tv + geom_point(color = "darkgreen", shape = 8, size = 4.8) + 
  theme(panel.grid = element_line(color = "lightgrey", linewidth = 0.8, linetype = 1), 
        panel.background = element_rect(color = "white", fill = "darkgrey")) +
  labs(title = "Happy TV", x ="Time TV", 
       y = "Happiness") #+ xlim(0,7) + ylim(45,95)
```
Exercise 7.2 Q3
TimeReading vs. TimeTV shows a strong negative correlation
TimeReading vs. Happiness shows a less strong, less negative correlation
TimeTV vs. Happiness shows a weak positive correlation

```{r covariance, echo = TRUE, include = TRUE}
# Find the covariance matrix
data_group <- cbind(survey$TimeReading, survey$TimeTV, survey$Happiness)
cov_data <- cov(data_group)
cov_data
```
Exercise 7.2 Q4
# TimeReading has a negative covariance with both TimeTV and Happiness. Happiness decreases half as fast as TimeTV as TimeReading increases

# As TimeTV increases Happiness increases rapidly as they have a relatively large positive covariance

```{r corellation, echo = TRUE, include = TRUE}
cor_data <- cor(data_group)
cor_data
```
Exercise 7.2 Q5
# TimeReading has a negative correlation with both TimeTV and Happiness. Happiness decreases half as fast as TimeTV as TimeReading increases

# As TimeTV increases Happiness increases rapidly as they have a relatively large positive correlation

# I think that correlation is better and easier to interpret for one primary reason. I believe it's better because there's a reference number other than zero i.e. it's bounded by 1 and -1

```{r corellation_2, echo = TRUE, include = TRUE}
# Find the corellation between TimeReading and TimeTV
cor_r_tv <- cor(survey$TimeReading, survey$TimeTV)
cor_r_tv
```
# Exercise 7.2 Q6
# TimeReading has a strong negative correlation with TimeTV. 

# We cannot assume causation from correlation. We cannot assume causation from correlation. We cannot assume causation from correlation. I would imagine we could still say that if you spend more time reading there's less time to spend watching TV, so it probably does have an effect.